### 2021-12-02 inst/MyArticle/Prolfqua.Rmd

---
journal: jacsat
type: article
author:
  - name: Witold E. Wolski
  - name: Christian Panse
  - name: Jonas Grossman
  - name: Paolo Nanni
  - name: Ralph Schlappbach
title:
  formatted: Prolfqua - \textsf{Pro}teomics label free quantification using linear models.
   class\footnote{A footnote for the title}
  short: prolfqua
abbr: IR,NMR,UV
keywords: American Chemical Society, \LaTeX
abstract: We introduce here the R package and benchmark linear modelling methods available in R for label free protein expression analysis.
bibliography: acstest.bib
csl: american-chemical-society.csl
output: rticles::acs_article
editor_options: 
  chunk_output_type: console
---

\begin{tocentry}
Some journals require a graphical entry for the Table of Contents.
This should be laid out ``print ready'' so that the sizing of the
text is correct.

Inside the \texttt{tocentry} environment, the font used is Helvetica
8\,pt, as required by \emph{Journal of the American Chemical
Society}.

The surrounding frame is 9\,cm by 3.5\,cm, which is the maximum
permitted for  \emph{Journal of the American Chemical Society}
graphical table of content entries. The box will not resize if the
content is too big: instead it will overflow the edge of the box.

This box and the associated title will always be printed on a
separate page at the end of the document.
\end{tocentry}

# Introduction

```{r setup, include=FALSE}
library(tidyverse)
library(prolfqua)
library(ggplot2)

knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

  The R-package for **pro**teomics **l**abel **f**ree **qua**ntification `prolfqua` (read: prolewka) evolved from functions and code snippets used to visualize and analyze label-free quantification data. To compute protein fold changes among treatment conditions, we first used t-test, linear models, or functions implemented in the package limma. We evaluated [MSStats]( 10.18129/B9.bioc.MSstats), [ROPECA](10.1038/s41598-017-05949-y) or [MSqRob](https://github.com/statOmics/MSqRob) all implemented in R, with the idea to integrate the various approaches in our analysis pipeline. Although these packages are implemented in R, model specification, input, and output formats differ widely and wildly. Furthermore, to better understand the inferences produced, we decided to use the R linear and mixed effect models directly and transparently and to reimplement some features from the other packages when considered useful. The R-package prolfqua is the outcome of this venture.

The _prolfqua_ package integrates most of the steps of the LFQ data analysis workflow: quality control, data normalization, protein aggregation, sample size estimation, modeling, and hypothesis testing. For instance, the quality control module makes it easy to determine the coefficients of variations (CV) for peptides and proteins within conditions, summarize and visualize missing data or intensities.

When developing _prolfqua_, we draw inspiration from packages such as _sf_, which uses data in a long tidy table format. We represent the data needed for analysis using a single data-frame in long format and an R6 configuration object. Using the configuration object, we specify what data is in which column, making it easy to integrate new inputs if provided in long tidy tables. All that is needed to incorporate Spectronaut, Skyline text output, or the __MSStats__ data format is a few code lines of code to update the configuration object. For software like MaxQuant or MSFragger, which writes the data into a wide table, with several intensity columns, one for each sample, we implemented methods that transform the data into a long tidy table. Relying on the long tidy data table enabled us to easily interface with many data manipulation, visualization, and modeling methods implemented in base R and the _tidyverse_.

 R linear model and linear mixed effect models allow modeling parallel designs, repeated measurements, factorial designs, and many more.  R's formula interface for linear models is flexible, widely used, and well documented. We integrate R's linear model and mixed model formula interface in _prolfqua_'s API. This glass box approach should make it easy to reimplement an analysis performed with prolfqua using base R or other programming languages by reading the analysis script and without looking at the package code. Knowledge of the R regression model infrastructure is an advantage when using our package. Acknowledging the formula interface's complexity and the popularity of MSstats, we provided functionality to derive the model formula from an MSstats data format.

The prolfqua package fits the linear model to all the proteins in the dataset. Afterward, we can compute contrast to test the hypothesis, perform ANOVA, or model selection using the likelihood ratio test for thousands of proteins. By exploiting the parallel structure of the analysis, _prolfqua_ implements t-statistic moderation to improve inference for small sample sizes [_limma_]. It also computes probabilities of differential protein regulation based on peptide level models [_ropeca_].

To optimize a data analysis workflow, prolfqua implements and integrates benchmarking methods, e.g., computing ROC curves. We implemented these methods to assess differences between linear models, mixed effect models, p-value moderation, or Bayesian regression models using the IonStar dataset.

We developed and improved the package by applying the "Eating your own dog food" principle. We used it to analyze SRM, PRM, and LFQ data and to generate analysis reports. The current iteration of the packages API groups the functionality into R6 classes (e.g., LFQPlotter, LFQSummarizer, LFQStats, Model, Contrast, etc.). We choose R6 for its visual debugging and code completion support in RStudio.

We believe that the package makes it easy to perform proteomics quantitative data analysis, generate visualizations, and integrate them into reproducible reports using Rmarkdown. You can install the package from [www.github.com/wolski/prolfqua](www.github.com/wolski/prolfqua), and we are working on making it available on Bioconductor.

# Methods

## Benchmarking
The benchmark dataset contains _H. sapiens_ proteins with constant concentrations and _E. coli_ proteins with varying concentration [@shen2018ionstar]. We know that for _H. sapiens_ proteins the difference $\beta$ between two dilutions should be $\beta = 0$ while for _E. coli_ proteins we know that the difference between dilutions should be $\beta \ne 0$.

We can use various statistics to examine this hypothesis: the contrast estimate $\beta$, or because we first log2 transformed the data the log2 fold-change (log2FC), the t-statistic $\frac{\beta}{\sqrt{var(\beta)}}$, the $p-value$  and moderated $p-value$ for models where the p-value can be obtained.

For Bayesian models we can use the posterior probabilities that $\beta=0$, e.g. the probability that $\beta < 0$ when the mean estimate $\hat{\beta}$ is greater $0$ or vice versa, i.e.  $P(\beta < 0| \hat{\beta} > 0)$ or $P(\beta > 0 | \hat{\beta} < 0)$, which we can compute from the posterior sample of the $\beta$ parameters. We can also use the statistic $\frac{mean(\beta)}{\sqrt{var(\beta)}}$ with $var{\beta}$ estimated from the posterior sample of the $\beta$ parameter.

For each statistic and each value of the statistics we then compute a confusion matrix (see Table \@ref(tab:confusionMatrix)).
From the confusion matrix we can compute measures such as true positive rate (TPR), false positive rate (FPR), false discovery proportion (FDP) or accuracy (ACC) which are given by:

```{r confusionMatrix}
table <- data.frame( c("beta != 0", "beta == 0", "Total"), matrix(c("TP","FP","R","FN","TN","","P","N","m"), ncol = 3, byrow = T))
colnames(table) <- c("Prediction \\ Truth","E.coli", "H.sapiens", "Total")
knitr::kable(table, caption = "Confusion matrix, TP - true positive, FP - false positive, FN - false negative, P - all positive cases (all E. coli proteins), N - all negative cases (all H. sapiens proteins), m- all proteins.", format="latex")

```

$$
\begin{aligned}
TPR &= \frac{TP}{TP+FN} = \frac{TP}{P}\\
FPR &= \frac{FP}{FP+TN} = \frac{FP}{N}\\
FDP &= \frac{FP}{TP + FP} = \frac{FP}{R}\\
ACC &= \frac{TP + TN}{m}
\end{aligned}
$$

By plotting the $TPR$ versus the $FPR$ we obtain the receiver operator characteristic curve (ROC curve). The area under the curve (AUC) or partial areas under the curve (pAUC), at various values of the $FPR$, are further measures of performance. By using these measures we can compare the performances of the statistics produced by the various methods examined.

A further question we can examine using the benchmark data is, how well the false discovery estimate (FDR) obtained to from the statistical model matches the false discovery proportion (FDP). The FDR is the expected value of the false discovery proportion. Ideally, the FDR should be an unbiased estimate of the $FDP$. By plotting the FDR gainst the FDP we can asses visually if these assumptions are met.

For the Bayesian models we obtain an FDR estimate by first computing the p-value using the statistic $\frac{mean(\beta)}{\sqrt{var(\beta)}}$ and assuming that it is normally distributed. We then obtain the FDR by adjusting the  p-values using the Benjamini-Hochberg procedure.

For some of the proteins, quantification results are missing in some conditions. All we required when attempting to fit a model for a protein is that there are at least two peptides per protein quantified in any of the twenty samples. That means that for some proteins, not all model parameters are estimable. 
In case of the linear model, the contrast function we implemented can handle partial models and provides estimates of some of the contrasts (see Table \@ref(tab:tablesMedpolish)). For the mixed effect models, the used `lmerTest::contest` function does not support models with missing parameter estimates, and hence we have 4 (all) or 0 (no) contrast estimates for a protein (see Table \@ref(tab:tablesMM)). For the Bayesian models obtained with _brms_ or _rstan_ we also only computed contrasts for models with complete fixed effect estimates (see Tables \@ref(tab:countsBrms), \@ref(tab:countsBrmsStudent) and \@ref(tab:countsRstand)). The ROPECA approach starts by fitting a model to each peptide and the contrast function can handle missing parameter estimates (see Table \@ref(tab:tablesRopeca)).
These differences result in different numbers of fold-change estimates for each type of model, i.e. in a different set of ground truths, which makes comparing model performances involved.
To deal with these differences, we computed our benchmark estimates, including the peptides and proteins, without fold-change estimates, by accordingly updating the numbers in the confusion matrix. To compute the $FPR$ and $TPR$ including the missing estimates we set $P =$ #all _E. coli_ peptides/proteins detected, and $N =$ #all _H. sapiens_ proteins detected.

## Benchmark Data preprocessing

We fitted the models using all the $20$ samples of the IonStar dataset (see Figure \@ref(fig:ionstar)), i.e. five different dilutions, with four technical replicates each. This allows to estimate ten different contrasts shown in Table \@ref(tab:setupContrasts). For benchmarking we only used the contrasts resulting in small fold-changes $\beta = 1.2,1.25,1.3(3),1.5$ listed in Table \@ref(tab:usedContrasts). Only those small fold-changes allowed us to see differences among the methods. For the larger fold-changes no differences can be seen among the models or statistics.

Data was preprocessed by $log_2$ transforming the peptide intensities, and subsequent robust z-score transformation. No other data filtering was used than the removal of _one hit wonders_, i.e. proteins with a single peptide assignment (For details see section [Datasets and Data preprocessing]). We did not use imputation.  The only differences we will observe are because of using different modelling methods.

$$
z = \frac{x - \bar{x}}{S} \sim \frac{x - \tilde{x}}{\tilde{S}}
$$  t fold change on the original scale, we multiply the z-score by the average variance of all the $N$ samples in the experiment.

$$
z' = z \times 1/N\sum_{i=1}^N S_i
$$
We decided to apply this simple transformation because it only requires to estimate two parameters per sample and works for experiments with thousands of proteins as well as for experiments where only a few hundreds of proteins per sample are measured.

```{r setupContrasts}

outpath <- "results_modelling_all"
#modelName  <- "Model"

DEBUG <- FALSE

Contrasts <- c(
  "dilution_(9/3)_3" =   "dilution.e - dilution.a",
  "dilution_(9/4.5)_2" =   "dilution.e - dilution.b",
  "dilution_(9/6)_1.5" =   "dilution.e - dilution.c",
  "dilution_(9/7.5)_1.2" =   "dilution.e - dilution.d",

  "dilution_(7.5/3)_2.5" =   "dilution.d - dilution.a",
  "dilution_(7.5/4.5)_1.6(6)" =   "dilution.d - dilution.b",
  "dilution_(7.5/6)_1.25" =   "dilution.d - dilution.c",

  "dilution_(6/3)_2" =   "dilution.c - dilution.a",
  "dilution_(6/4.5)_1.3(3)" =   "dilution.c - dilution.b",

  "dilution_(4.5/3)_1.5" =   "dilution.b - dilution.a"
)

tt <- data.frame(contrastsName = names(Contrasts),  contrasts = Contrasts)
tt <- tt %>% separate(contrastsName, c("dilution", "ratio", "expected fold change"), sep =  "_", remove = FALSE)

tt
```

```{r usedContrasts}
relevantContrasts <- c("dilution_(9/7.5)_1.2","dilution_(7.5/6)_1.25", "dilution_(6/4.5)_1.3(3)", "dilution_(4.5/3)_1.5" )

tt <- tt %>% dplyr::filter(contrastsName  %in% relevantContrasts)
tt <- select(tt, all_of(c("contrasts","ratio", "expected fold change")))
knitr::kable(tt, caption = "Contrasts used for benchmark.", format = "latex")

```

Furthermore, we will only model protein fold-changes of proteins were at least two peptides were observed. The reason is that peptide identification has an associated error which is significantly reduced if two independent peptide identifications of a protein are observed. Many proteomics journals require to only report proteins identified by two peptides. Proteins identified by only a single peptide are sometimes called "one-hit wonders" [@mendoza2018flexible].

In order to remove systematic differences among samples, peptide intensities need to be transformed and scaled. The transformation aims to remove heteroscedasticity while the scaling aims to remove systematic differences among samples. Because the size of the error is proportional to the intensity, by $\log$ transforming the intensities this error can be modelled with $\epsilon \propto N(0, \sigma^2)$.

Valikangas [@valikangas2016systematic] and colleagues discuss and benchmark various methods of peptide or protein intensity normalization such as variance stabilizing normalization [@huber2002variance] or quantile normalization [@bolstad2003comparison]. In this work we will be using a robust version of the z-score, where instead of the mean we use the median and instead of the standard deviation the median absolute deviation (mad):

$$
z = \frac{x - \bar{x}}{S} \sim \frac{x - \tilde{x}}{\tilde{S}}
$$

Because we need to estimate the protein fold-changes on the original scale, we have to multiply the $z$-score by the average variance of all the $N$ samples in the experiment.

$$
z' = z \times 1/N\sum_{i=1}^N S_i
$$
We decided to apply this simple transformation because it only requires to estimate two parameters per sample and works for experiments with thousands of proteins as well as for experiments where only a few hundreds of proteins per sample are measured.
For the Ionstar dataset we used the intensities of _H. sapiens_ proteins, whose concentrations do not change, to determine $\bar{x}$ and $S$ and than applied it to all the intensities (including _E. coli_) in the sample.

Figure \@ref(fig:scaling) shows the distribution of the peptide intensities within the samples before and after the intensity scaling. Figure \@ref(fig:variancestable) shows the coefficient of variation before log transforming and scaling the data and the distribution of the standard deviations afterwards.

(ref:scaling) Density plot of peptide intensities before (panel A) and after (panel B) data transformation and scaling.

```{r scaling, fig.cap="(ref:scaling)", fig.width=7, fig.height=7}
is <- prolfqua_data('data_ionstar')
pepsum <- is$filtered()
summarised <- is$subset_normalized()

p1 <-
  prolfqua::plot_intensity_distribution_density(pepsum$data,
                                                  pepsum$config) +
  theme(legend.position = "none") +
  labs(tag = "A") + theme(legend.position = "none")
p2 <-
  prolfqua::plot_intensity_distribution_density(
    summarised$data,
    summarised$config) +
  labs(tag = "B") + theme(legend.position = "none")
rocp <-
      ggpubr::ggarrange(
        p1,
        p2,
        nrow = 1,
        common.legend = TRUE,
        legend = "none"
      )
rocp
```

## Linear model

Here we show how the two-stage modelling by first estimating the protein intensities using the Tukey median polish and then fitting a linear model performs.

# Results and discussion

In this section we compare the results for the various models implemented and discussed previously, namely:

* (`prot_med_lm`) : linear model fitted to protein estimates obtained using Tukey's median polish 
* (`prot_lme4`) : mixed effects model fitted using _lme4_ 
* (`prot_ROPECA`) : linear models fitted on peptide level and aggregation of moderated p-values using the beta distribution
* (`prot_rstan`) : mixed effects model fitted with _rstan_ and with prior over $\varepsilon$ 
* (`prot_brms`) : mixed effects model fitted with _brms_ and default priors 
* (`prot_brms_student_t`) : mixed effects model fitted with _brms_ and shrinkage prior over fixed effects coefficients .

A relevant parameter is the number of proteins for which we estimated the contrasts (see Table \@ref(tab:completeCasesTab)). It indicates how robust the models are in the presence of missing data. We observe, that when using the linear model on protein level intensities or the bayesian version of the mixed effects models we were able to estimate for 4024 (out of the 4099) proteins all the four examined contrasts (complete cases). When using the mixed effect model implementation in _lme4_ this number drops to $4013$, while when summarizing peptide level models it drops further to $3999$ proteins.

The different number of proteins for which we obtained statistics of contrasts makes comparing the scores for various models complicated because the set of proteins differs, e.g. $4024$ versus $3999$. 
Therefore, we incorporate the failure to generate statistics for contrasts into the benchmark by setting $P$ and $N$ equal to the number of all proteins (see Table \@ref(tab:confusionMatrix)) and then computing the $FPR$ and $TPR$.

## Outline

# Extra information when writing JACS Communications

\begin{acknowledgement}

The authors thank technology platform fund of the University of Zurich.

\end{acknowledgement}

\begin{suppinfo}

This will usually read something like: ``Experimental procedures and
characterization data for all new compounds. The class will
automatically add a sentence pointing to the information on-line:

\end{suppinfo}
## References