### 2022-10-22 inst/MyArticle/prolfqua_supplement.Rmd
updated supplement.

---
title: "Supporting Information:\n prolfqua - A Comprehensive R-package for Proteomics Differential Expression Analysis"
author: 
  - name: Witold E. Wolski
    affiliation:
      - Functional Genomics Center Zurich (FGCZ), ETH Zurich / University of Zurich, Winterthurerstrasse 190, 8057 Zurich, Switzerland
      - Swiss Institute of Bioinformatics (SIB), Quartier Sorge - Batiment Amphipole, 1015 Lausanne, Switzerland
    email: wew@fgcz.ethz.ethz.ch
  - name: Paolo Nanni
    affiliation: Functional Genomics Center Zurich (FGCZ), ETH Zurich / University of Zurich, Winterthurerstrasse 190, 8057 Zurich, Switzerland
  - name: Jonas Grossmann
    affiliation:
      - Functional Genomics Center Zurich (FGCZ), ETH Zurich / University of Zurich, Winterthurerstrasse 190, 8057 Zurich, Switzerland
      - Swiss Institute of Bioinformatics (SIB), Quartier Sorge - Batiment Amphipole, 1015 Lausanne, Switzerland
  - name: Maria d'Errico
    affiliation:
      - Functional Genomics Center Zurich (FGCZ), ETH Zurich / University of Zurich, Winterthurerstrasse 190, 8057 Zurich, Switzerland
      - Swiss Institute of Bioinformatics (SIB), Quartier Sorge - Batiment Amphipole, 1015 Lausanne, Switzerland
  - name: Ralph Schlapbach
    affiliation: Functional Genomics Center Zurich (FGCZ), ETH Zurich / University of Zurich, Winterthurerstrasse 190, 8057 Zurich, Switzerland
  - name: Christian Panse
    affiliation:
      - Functional Genomics Center Zurich (FGCZ), ETH Zurich / University of Zurich, Winterthurerstrasse 190, 8057 Zurich, Switzerland
      - Swiss Institute of Bioinformatics (SIB), Quartier Sorge - Batiment Amphipole, 1015 Lausanne, Switzerland
package: prolfqua
output:
  BiocStyle::pdf_document
abstract: |
  This document contains the supplements for the prolfqua manuscript available
  through \url{https://doi.org/10.1101/2022.06.07.494524}.
vignette: |
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
#bibliography: prolfqua.bib
---

\renewcommand{\thepage}{S--\arabic{page}}
\renewcommand{\tablename}{Supp. Table}
\renewcommand{\figurename}{Supp. Figure}

\newpage

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, error = FALSE, warning = FALSE, message = FALSE)
```

# prolfqua Installation

To install the `r BiocStyle::Githubpkg("fgcz/prolfqua")` package ([ACS_JPR](https://github.com/fgcz/prolfqua/releases/tag/v0.4.3) release), including all R package dependencies, execute in the R console:

```{r installation, echo=TRUE, eval=FALSE}
remotes::install_github("https://github.com/fgcz/prolfqua/releases/tag/v0.4.5",
    dependencies = TRUE, build_vignettes=TRUE)
```

The R-markdown files with all the R code to run the benchmark for `r BiocStyle::Githubpkg("prolfqua")`, `r BiocStyle::Biocpkg("msqrob2")`, `BiocStyle::Biocpkg("MSStats")` and `BiocStyle::Biocpkg("proDA")` are available in the R package `r BiocStyle::Githubpkg('wolski/prolfquabenchmark')`.

```{r installation2, echo=TRUE, eval=FALSE}
remotes::install_gitlab("wolski/prolfquadata", host="gitlab.bfabric.org")
remotes::install_github("https://github.com/wolski/prolfquabenchmark",
    dependencies = TRUE, build_vignettes=TRUE)
```

# Comparing partial areas under the curve at 0.1 FDR for Ionstar/MaxQuant dataset

Here we test if we should reject the null hypothesis that the area under the ROC curve at 0.1 FDR does not differ for 
msqrob2, proDA, and prolfqua when computed based on scaled p.values. To this task we use the function roc.test from the R package `r BiocStyle::CRANpkg("pROC")`.

```{r roctest, include = FALSE, eval = TRUE}
allBenchmarks <- readRDS("allBenchmarks.RDS")
ttmsqrob <- allBenchmarks$benchmark_msqrob$data()
ddmsqrob <- pROC::roc(ttmsqrob$TP, ttmsqrob$scaled.p.value, partial.auc = c(1, 0.9))

ttprolfqua <- allBenchmarks$benchmark_merged$data()
ddprolfqua <- pROC::roc(ttprolfqua$TP, ttprolfqua$scaled.p.value, partial.auc = c(1, 0.9))
ttproda <- allBenchmarks$benchmark_proDA$data()
ddproda <- pROC::roc(ttproda$TP, ttproda$scaled.p.value, partial.auc = c(1, 0.9))

tmp <- c(msqrob2_vs_prolfqua = pROC::roc.test(ddmsqrob,ddprolfqua, progress = "none")$p.value,
         msqrob2_vs_proda = pROC::roc.test(ddmsqrob,ddproda, progress = "none")$p.value,
         prolfqua_vs_proda = pROC::roc.test(ddprolfqua,ddproda, progress = "none")$p.value)

```

The Table \@ref(tab:shwoROCtestREsultsIonStarMQ) shows the p-values of the bootstrap test for two ROC curves. We observe that for this benchmark data, there are no significant differences among the $AUC_{10}$ for the three packages.

```{r shwoROCtestREsultsIonStarMQ}
cap = "p-values for pairwise comparsions of partial AUC at 0.1 for msqrob2, proDA and prolfqua for IonStar MaxQuant data."
knitr::kable(tibble(pROC_test = names(tmp), p.value = round(tmp, digits = 2)), format = "latex", caption = cap)

```

# Benchmarking msqrob2, proDA and prolfqua using the CPTAC dataset

The Rmarkdown file which created these results, with more details about the CPTAC / MaxQuant dataset, can be found here:
["Benchmarking of msqrob2, prolfqua and proDA using the CPTAC dataset MaxQuant output"](https://wolski.github.io/prolfquabenchmark/articles/Benchmark_cptac.html)

```{r}
getpath <- function(filN){
    f1 <- paste0("../../inst/Benchresults/",filN)
    if (f1 == "") {
        f1 <- system.file("Benchresults",filN, package = "prolfquabenchmark")
    }
    message(f1)
    return(f1)
}
CPTACBenchmarks <- readRDS(getpath("CPTAC_Benchmark.RDS"))
mrob <- CPTACBenchmarks$benchmark_msqrob$pAUC_summaries()$ftable$content
mrob$package <- "msqrob2"
proda <- CPTACBenchmarks$benchmark_proDA$pAUC_summaries()$ftable$content
proda$package <- "proDA"
prolfqua <- CPTACBenchmarks$benchmark_merged$pAUC_summaries()$ftable$content
prolfqua$package <- "prolfqua"

all <- dplyr::bind_rows(list(mrob, proda, prolfqua))
all <- all |> filter(contrast == "b_vs_a")
all$what[all$what == "statistic"] <- "t_statistic"
all$what[all$what == "t"] <- "t_statistic"
all$what[all$what == "scaled.pval"] <- "scaled.p.value"
all$what[all$what == "logFC"] <- "diff"

```

Figure \@ref(fig:plotCPTACbarplot) shows the partial area under the curve for `r BiocStyle::Biocpkg("msqrob2")`, `r BiocStyle::Biocpkg("proDA")` and `r BiocStyle::Githubpkg("fgcz/prolfqua")`.

```{r plotCPTACbarplot, fig.cap= "Barplot showing the partial area under the ROC at 0.1 FDR."}
all |> ggplot(aes(x = what, y = pAUC_10, group = package, color = NULL, fill = package)) +
  geom_bar(stat = "identity",  position = position_dodge())

```

```{r roctestCPTAC, include = TRUE, message=FALSE, error=FALSE}

ttmsqrob <- CPTACBenchmarks$benchmark_msqrob$data()
ddmsqrob <- pROC::roc(ttmsqrob$TP, ttmsqrob$scaled.pval, partial.auc = c(1, 0.9))
ttprolfqua <- CPTACBenchmarks$benchmark_merged$data()
ddprolfqua <- pROC::roc(ttprolfqua$TP, ttprolfqua$scaled.p.value, partial.auc = c(1, 0.9))
ttproda <- CPTACBenchmarks$benchmark_proDA$data()
ddproda <- pROC::roc(ttproda$TP, ttproda$scaled.pval, partial.auc = c(1, 0.9))

tmp <- c(msqrob2_vs_prolfqua = pROC::roc.test(ddmsqrob,ddprolfqua, progress = "none")$p.value,
         msqrob2_vs_proda = pROC::roc.test(ddmsqrob,ddproda, progress = "none")$p.value,
         prolfqua_vs_proda = pROC::roc.test(ddprolfqua,ddproda, progress = "none")$p.value)

```

Table \@ref(tab:shwoROCtestREsultsCPTAC) shows the results of the Bootstrap test for two ROC curves. 
We observe that for this benchmark data, there are no significant differences among the $AUC_{10}$ for the three packages.

```{r shwoROCtestREsultsCPTAC}
cap = "p-values for pairwise comparsions of partial AUC for msqrob2, proDA and prolfqua CPTAC/MaxQuant data."
knitr::kable(tibble(pROC_test = names(tmp), p.value = round(tmp, digits = 2)), format = "latex", caption = cap)

```

# Benchmark IonQuant FragPipe v15 combined_protein.tsv

The Rmarkdown file which created these results, with more details about the IonQuant FragPipe v15 dataset, can be found here:
["Benchmarking proDA and prolfqua using Ionstar Dataset and FragPipe v15 compbined_protein output"](https://wolski.github.io/prolfquabenchmark/articles/Benchmark_cptac.html)

Direct comparison with `msqrob2` is impossible since, to fit the dropout model, the peptide intensities are required, while here, we are starting the analysis from the `combined_proteins.tsv` file.

```{r FragPipe}
FragPipeBenchmarks <- readRDS(getpath("FragPipev15_comb_prot.RDS"))
proda <- FragPipeBenchmarks$benchmark_proDA$pAUC_summaries()$ftable$content
proda$package <- "proDA"
prolfqua <- FragPipeBenchmarks$benchmark_prolfqua$pAUC_summaries()$ftable$content
prolfqua$package <- "prolfqua"

tmp <- dplyr::bind_rows(proda, prolfqua)
tmp$what |> unique()
tmp$what[tmp$what == "statistic"] <- "t_statistic"
tmp$what[tmp$what == "scaled.pval"] <- "scaled.p.value"
tmp <- tmp |> filter(contrast == "all")
```

Figure \@ref(fig:plotCPTACbarplot) shows the partial area under the curve for `r BiocStyle::Biocpkg("msqrob2")`, `r BiocStyle::Biocpkg("proDA")` and `r BiocStyle::Githubpkg("fgcz/prolfqua")`.

```{r plotFragPipebarplot, fig.cap= "Barplot showing the partial area under the ROC at 0.1 FDR."}
tmp |> ggplot2::ggplot(ggplot2::aes(x = what, y = pAUC_10, group = package, color = NULL, fill = package)) +
  ggplot2::geom_bar(stat = "identity",  position = ggplot2::position_dodge()) +
  ggplot2::facet_wrap(~ contrast) +
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 90, vjust = 0.5, hjust = 1))

```

```{r roctestFRAGPIPE, include = TRUE, message=FALSE, error=FALSE}

ttprolfqua <- FragPipeBenchmarks$benchmark_prolfqua$data()
ddprolfqua <- pROC::roc(ttprolfqua$TP, ttprolfqua$scaled.p.value, partial.auc = c(1, 0.9))
ttproda <- FragPipeBenchmarks$benchmark_proDA$data()
ddproda <- pROC::roc(ttproda$TP, ttproda$scaled.pval, partial.auc = c(1, 0.9))

tmp <- c(prolfqua_vs_proda = pROC::roc.test(ddprolfqua,ddproda, progress = "none")$p.value)

```

Table \@ref(tab:shwoROCtestREsultsFRAGPIPE) shows the results of the Bootstrap test for the difference of two ROC curves. We observe that for this benchmark data, there are no significant differences among the $AUC_{10}$ for the two packages.

```{r shwoROCtestREsultsFRAGPIPE}
cap = "p-values for pairwise comparsions of partial AUC for msqrob2, proDA and prolfqua CPTAC/MaxQuant data."
knitr::kable(tibble(pROC_test = names(tmp), p.value = round(tmp, digits = 2)), format = "latex", caption = cap)

```

# Benchmark Vignettes

Pre-build version of the vignettes are available:

- [Benchmarking of methods implemented in prolfqua using the IonStar dataset MaxQuant](https://wolski.github.io/prolfquabenchmark/articles/BenchmarkingIonstarData.html) 
- [Benchmarking of MSstats using the IonStar dataset MaxQuant](https://wolski.github.io/prolfquabenchmark/articles/Benchmark_MSStats.html) 
- [Benchmarking of proDA  using the IonStar dataset MaxQuant](https://wolski.github.io/prolfquabenchmark/articles/Benchmark_proDA_medpolish.html) 
- [Benchmarking of msqrob2  using the IonStar dataset MaxQuant](https://wolski.github.io/prolfquabenchmark/articles/BenchmarkMSqRob2.html)  
- [Benchmarking of proDA and proflqua using IonStar dataset MSFragger IonQuant combined_protein.txt](https://wolski.github.io/prolfquabenchmark/articles/BenchmarkMSFraggerProteinIonStar.html)

# Creating a prolfqua configuration

The following code demonstrates how we use `r BiocStyle::Githubpkg("fgcz/prolfqua")` to analyze protein intensities reported in the MSFragger `combined_protein.tsv` file.
First, we create a tidy table containing the protein abundances by reading the `combined_protein.tsv` file using  `tidy_MSFragger_combined_protein.` Then, we read the sample annotation from the file `annotation.xlsx` file. Next, we create an `AnalysisTableAnnotation` R6 object.
Bottom-up proteomics data is hierarchical, i.e., a protein has peptides, peptides might be modified, etc. Therefore, the `AnalysisTableAnnotation` has a `hierarchy` field storing a list with an entry for each hierarchy level.
Since `combined_portein.tsv` only holds protein level data, the hierarchy list has one element, and we use it to specify which column contains the protein identifiers. We also need to define which column contains the protein abundances we want to use for the data analysis.
Finally, we have to specify which columns contain the explanatory variables of the analysis. The `AnalysisTableAnnotation` has the field `factors,` a list with as many entries as explanatory variables. Here we include two explanatory variables, the dilution, specified in the column 'sample', and 'run' stored in the column 'run_ID', representing the order of the measurement.

```{r echo=TRUE}
datadir <- file.path(find.package("prolfquadata") , "quantdata")
inputFragfile <-  file.path(datadir, "MSFragger_IonStar2018_PXD003881.zip")
inputAnnotation <- file.path(datadir, "annotation_Ionstar2018_PXD003881.xlsx")
# read input annotation
annotation <- readxl::read_xlsx(inputAnnotation)

protein <- tibble::as_tibble(
    read.csv(unz(inputFragfile,"IonstarWithMSFragger/combined_protein.tsv"),
             header = TRUE, sep = "\t", stringsAsFactors = FALSE))

# read combined_protein.tsv 
protein <- prolfqua::tidy_FragPipe_combined_protein_deprec(protein)
# remove proteins identified by a single peptide
protein <- protein |> dplyr::filter(unique.stripped.peptides > 1)

# annotate the data
merged <- dplyr::inner_join(annotation, protein)
atable <- prolfqua::AnalysisTableAnnotation$new()
atable$fileName = "raw.file"
# specify column containing protein identifiers
atable$hierarchy[["protein_Id"]] = "protein"

# column with protein abundances
atable$set_response("total.intensity")

# the factors of the analysis
atable$factors[["dilution."]] = "sample"
atable$factors[["run"]] = "run_ID"

config <- prolfqua::AnalysisConfiguration$new(atable)

adata <- prolfqua::setup_analysis(merged, config)
lfqdata <- prolfqua::LFQData$new(adata, config)
# show number of proteins in the dataset
lfqdata$hierarchy_counts()

```

# Miscellaneous

(ref:tabCompletion) The screenshot displays the command-line completion (tab completion) of RStudio on the `prolfqua::LFQData` R6 object. In the example, it shows the getter methods of the object.

```{r tabCompletion, echo=FALSE, fig.cap="(ref:tabCompletion)", out.width = '66%'}
knitr::include_graphics("graphics/codeSnippet1TabCompletion.png")
```

```{r proLFQuaSticker, echo=TRUE, out.height="5cm", eval=TRUE, fig.cap="Sticker maintainer: Witold E. Wolski; License: Creative Commons Attribution CC-BY. Feel free to share and adapt, but don't forget to credit the author."}
file.path("graphics/hexStickerProlfqua.png") |>
  knitr::include_graphics()
```

# Session info {.unnumbered}

```{r sessionInfo, echo=FALSE}
pander::pander(sessionInfo())
```