### 2022-10-26 inst/MyArticle/prolfqua_supplement.Rmd
explicit dplyr::filter etc. close #24

---
title: "Supporting Information:\n prolfqua - A Comprehensive R-package for Proteomics Differential Expression Analysis"
author: 
  - name: Witold E. Wolski
    affiliation:
      - Functional Genomics Center Zurich (FGCZ), ETH Zurich / University of Zurich, Winterthurerstrasse 190, 8057 Zurich, Switzerland
      - Swiss Institute of Bioinformatics (SIB), Quartier Sorge - Batiment Amphipole, 1015 Lausanne, Switzerland
    email: wew@fgcz.ethz.ethz.ch
  - name: Paolo Nanni
    affiliation: Functional Genomics Center Zurich (FGCZ), ETH Zurich / University of Zurich, Winterthurerstrasse 190, 8057 Zurich, Switzerland
  - name: Jonas Grossmann
    affiliation:
      - Functional Genomics Center Zurich (FGCZ), ETH Zurich / University of Zurich, Winterthurerstrasse 190, 8057 Zurich, Switzerland
      - Swiss Institute of Bioinformatics (SIB), Quartier Sorge - Batiment Amphipole, 1015 Lausanne, Switzerland
  - name: Maria d'Errico
    affiliation:
      - Functional Genomics Center Zurich (FGCZ), ETH Zurich / University of Zurich, Winterthurerstrasse 190, 8057 Zurich, Switzerland
      - Swiss Institute of Bioinformatics (SIB), Quartier Sorge - Batiment Amphipole, 1015 Lausanne, Switzerland
  - name: Ralph Schlapbach
    affiliation: Functional Genomics Center Zurich (FGCZ), ETH Zurich / University of Zurich, Winterthurerstrasse 190, 8057 Zurich, Switzerland
  - name: Christian Panse
    affiliation:
      - Functional Genomics Center Zurich (FGCZ), ETH Zurich / University of Zurich, Winterthurerstrasse 190, 8057 Zurich, Switzerland
      - Swiss Institute of Bioinformatics (SIB), Quartier Sorge - Batiment Amphipole, 1015 Lausanne, Switzerland
package: prolfqua
output:
  BiocStyle::pdf_document
abstract: |
  This document contains the supplements for the prolfqua manuscript available
  through \url{https://doi.org/10.1101/2022.06.07.494524}.
vignette: |
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
#bibliography: prolfqua.bib
---

\renewcommand{\thepage}{S--\arabic{page}}
\renewcommand{\tablename}{Supp. Table}
\renewcommand{\figurename}{Supp. Figure}

\newpage

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, error = FALSE, warning = FALSE, message = FALSE)
```

The sourcode used to produce this Document can be found in the  [prolfqua_supplement.Rmd](https://github.com/wolski/prolfquabenchmark/blob/main/inst/MyArticle/prolfqua_supplement.Rmd) file.

# prolfqua Installation

To install the `r BiocStyle::Githubpkg("fgcz/prolfqua")` package ([ACS_JPR](https://github.com/fgcz/prolfqua/releases/tag/v0.4.3) release), including all R package dependencies, execute in the R console:

```{r installation, echo=TRUE, eval=FALSE}
remotes::install_github("https://github.com/fgcz/prolfqua/releases/tag/v0.4.5",
    dependencies = TRUE, build_vignettes=TRUE)
```

The R-markdown files with all the R code to run the benchmark for `r BiocStyle::Githubpkg("prolfqua")`, `r BiocStyle::Biocpkg("msqrob2")`, `BiocStyle::Biocpkg("MSStats")` and `BiocStyle::Biocpkg("proDA")` are available in the R package `r BiocStyle::Githubpkg('wolski/prolfquabenchmark')`.

```{r installation2, echo=TRUE, eval=FALSE}
remotes::install_gitlab("wolski/prolfquadata", host="gitlab.bfabric.org")
remotes::install_github("https://github.com/wolski/prolfquabenchmark",
    dependencies = TRUE, build_vignettes=TRUE)
```

# Comparing partial areas under the curve at 0.1 FDR for Ionstar/MaxQuant dataset

We test if we should reject the null hypothesis that the area under the ROC curve at 0.1 FDR, computed based on scaled p.values, does not differ for msqrob2, proDA, and prolfqua (Figure 5 in the article). To this task we use the function `roc.test` from the R package `r BiocStyle::CRANpkg("pROC")`.

```{r roctest, include = TRUE, eval = TRUE}
allBenchmarks <- readRDS("allBenchmarks.RDS")
ttmsqrob <- allBenchmarks$benchmark_msqrob$data()
ddmsqrob <- pROC::roc(ttmsqrob$TP, ttmsqrob$scaled.p.value, partial.auc = c(1, 0.9))

ttprolfqua <- allBenchmarks$benchmark_merged$data()
ddprolfqua <- pROC::roc(ttprolfqua$TP, ttprolfqua$scaled.p.value, partial.auc = c(1, 0.9))
ttproda <- allBenchmarks$benchmark_proDA$data()
ddproda <- pROC::roc(ttproda$TP, ttproda$scaled.p.value, partial.auc = c(1, 0.9))

tmp <- c(msqrob2_vs_prolfqua = pROC::roc.test(ddmsqrob,ddprolfqua, progress = "none")$p.value,
         msqrob2_vs_proda = pROC::roc.test(ddmsqrob,ddproda, progress = "none")$p.value,
         prolfqua_vs_proda = pROC::roc.test(ddprolfqua,ddproda, progress = "none")$p.value)

```

The Table \@ref(tab:shwoROCtestREsultsIonStarMQ) shows the p-values of the pairwise bootstrap test for partial areas under the ROC curves. We will not reject the null hypothesis that there are no difference among the ROC curves.

```{r shwoROCtestREsultsIonStarMQ}
cap = "p-values for pairwise comparsions of partial AUC at 0.1 for msqrob2, proDA and prolfqua for IonStar MaxQuant data."
knitr::kable(tibble::tibble(pROC_test = names(tmp), p.value = round(tmp, digits = 2)), format = "latex", caption = cap)

```

# Benchmarking msqrob2, proDA and prolfqua using the CPTAC dataset

We present here the results of benchmarking the msqrob2, proDA and prolfqua packages using the CPTAC dataset.
The Rmarkdown file which generated these results, with more details about the CPTAC / MaxQuant dataset, can be found here:
["Benchmarking of msqrob2, prolfqua and proDA using the CPTAC dataset MaxQuant output"](https://wolski.github.io/prolfquabenchmark/articles/Benchmark_cptac.html)

```{r}
getpath <- function(filN){
    f1 <- paste0("../../inst/Benchresults/",filN)
    if (f1 == "") {
        f1 <- system.file("Benchresults",filN, package = "prolfquabenchmark")
    }
    message(f1)
    return(f1)
}
CPTACBenchmarks <- readRDS(getpath("CPTAC_Benchmark.RDS"))
mrob <- CPTACBenchmarks$benchmark_msqrob$pAUC_summaries()$ftable$content
mrob$package <- "msqrob2"
proda <- CPTACBenchmarks$benchmark_proDA$pAUC_summaries()$ftable$content
proda$package <- "proDA"
prolfqua <- CPTACBenchmarks$benchmark_merged$pAUC_summaries()$ftable$content
prolfqua$package <- "prolfqua"

all <- dplyr::bind_rows(list(mrob, proda, prolfqua))
all <- all |> 
  dplyr::filter(contrast == "b_vs_a")
all$what[all$what == "statistic"] <- "t_statistic"
all$what[all$what == "t"] <- "t_statistic"
all$what[all$what == "scaled.pval"] <- "scaled.p.value"
all$what[all$what == "logFC"] <- "diff"

```

Figure \@ref(fig:plotCPTACbarplot) shows the partial area under the curve for `r BiocStyle::Biocpkg("msqrob2")`, `r BiocStyle::Biocpkg("proDA")` and `r BiocStyle::Githubpkg("fgcz/prolfqua")`, when using the difference among the groups, the scaled p-values or the t-statistics to compute the ROC curves.

```{r plotCPTACbarplot, fig.cap= "Barplot showing the partial area under the ROC at 0.1 FDR."}
all |> ggplot2::ggplot(ggplot2::aes(x = what, y = pAUC_10, group = package, color = NULL, fill = package)) +
  ggplot2::geom_bar(stat = "identity",  position = ggplot2::position_dodge())

```

```{r roctestCPTAC, include = TRUE, message=FALSE, error=FALSE}

ttmsqrob <- CPTACBenchmarks$benchmark_msqrob$data()
ddmsqrob <- pROC::roc(ttmsqrob$TP, ttmsqrob$scaled.pval, partial.auc = c(1, 0.9))
ttprolfqua <- CPTACBenchmarks$benchmark_merged$data()
ddprolfqua <- pROC::roc(ttprolfqua$TP, ttprolfqua$scaled.p.value, partial.auc = c(1, 0.9))
ttproda <- CPTACBenchmarks$benchmark_proDA$data()
ddproda <- pROC::roc(ttproda$TP, ttproda$scaled.pval, partial.auc = c(1, 0.9))

tmp <- c(msqrob2_vs_prolfqua = pROC::roc.test(ddmsqrob,ddprolfqua, progress = "none")$p.value,
         msqrob2_vs_proda = pROC::roc.test(ddmsqrob,ddproda, progress = "none")$p.value,
         prolfqua_vs_proda = pROC::roc.test(ddprolfqua,ddproda, progress = "none")$p.value)

```

Table \@ref(tab:shwoROCtestREsultsCPTAC) shows the results of the Bootstrap test for two ROC curves. 
We observe that for this benchmark data, we can not reject the null hypothesis that there are no significant differences among the $pAUC_{10}$ (ROC based on scaled p-value) for the three packages.

```{r shwoROCtestREsultsCPTAC}
cap = "p-values for pairwise comparsions of partial AUC for msqrob2, proDA and prolfqua CPTAC/MaxQuant data."
knitr::kable(tibble::tibble(pROC_test = names(tmp), p.value = round(tmp, digits = 2)), format = "latex", caption = cap)

```

# Benchmark based on IonStar FragPipe v15 combined_protein.tsv

The Rmarkdown file which created these results, with more details about the IonStar FragPipe v15 dataset, can be found here:
["Benchmarking proDA and prolfqua using Ionstar Dataset and FragPipe v15 compbined_protein output"](https://wolski.github.io/prolfquabenchmark/articles/Benchmark_cptac.html)

Direct comparison with `msqrob2` is impossible since to fit the hurdle model, the peptide intensities are required, while here, we are starting the analysis from the `combined_proteins.tsv` file. Starting from protein level intensities has the advantage that the effect size estimates obtained from the model agree with the input data, a feature frequently requested by the biologist.

```{r FragPipe}
FragPipeBenchmarks <- readRDS(getpath("FragPipev15_comb_prot.RDS"))
proda <- FragPipeBenchmarks$benchmark_proDA$pAUC_summaries()$ftable$content
proda$package <- "proDA"
prolfqua <- FragPipeBenchmarks$benchmark_prolfqua$pAUC_summaries()$ftable$content
prolfqua$package <- "prolfqua"

tmp <- dplyr::bind_rows(proda, prolfqua)
tmp$what[tmp$what == "statistic"] <- "t_statistic"
tmp$what[tmp$what == "scaled.pval"] <- "scaled.p.value"
tmp <- tmp |> 
  dplyr::filter(contrast == "all")
```

Figure \@ref(fig:plotCPTACbarplot) shows the partial area under the curve for `r BiocStyle::Biocpkg("msqrob2")`, `r BiocStyle::Biocpkg("proDA")` and `r BiocStyle::Githubpkg("fgcz/prolfqua")`, computed using the effect size estimate among groups (diff), the scaled p-value (scaled.p.value) and the t-statistics (t_statistics),

```{r plotFragPipebarplot, fig.cap= "Barplot showing the partial area under the ROC at 0.1 FDR."}
tmp |> ggplot2::ggplot(ggplot2::aes(x = what, y = pAUC_10, group = package, color = NULL, fill = package)) +
  ggplot2::geom_bar(stat = "identity",  position = ggplot2::position_dodge()) +
  ggplot2::facet_wrap(~ contrast) +
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 90, vjust = 0.5, hjust = 1))

```

```{r roctestFRAGPIPE, include = TRUE, message=FALSE, error=FALSE}

ttprolfqua <- FragPipeBenchmarks$benchmark_prolfqua$data()
ddprolfqua <- pROC::roc(ttprolfqua$TP, ttprolfqua$scaled.p.value, partial.auc = c(1, 0.9))
ttproda <- FragPipeBenchmarks$benchmark_proDA$data()
ddproda <- pROC::roc(ttproda$TP, ttproda$scaled.pval, partial.auc = c(1, 0.9))

tmp <- c(prolfqua_vs_proda = pROC::roc.test(ddprolfqua,ddproda, progress = "none")$p.value)

```

Table \@ref(tab:shwoROCtestREsultsFRAGPIPE) shows the results of the Bootstrap test for the difference of two ROC curves. We observe that for this benchmark data, there are no significant differences among the $pAUC_{10}$ (for scaled p-value) for the two packages.

```{r shwoROCtestREsultsFRAGPIPE}
cap = "p-values for pairwise comparsions of partial AUC for proDA and prolfqua CPTAC/MaxQuant data."
knitr::kable(tibble::tibble(pROC_test = names(tmp), p.value = round(tmp, digits = 2)), format = "latex", caption = cap)

```

# Benchmark Vignettes

Pre-build version of the vignettes are available:

- [Benchmarking of methods implemented in prolfqua using the IonStar dataset MaxQuant](https://wolski.github.io/prolfquabenchmark/articles/BenchmarkingIonstarData.html) 
- [Benchmarking of MSstats using the IonStar dataset MaxQuant](https://wolski.github.io/prolfquabenchmark/articles/Benchmark_MSStats.html) 
- [Benchmarking of proDA  using the IonStar dataset MaxQuant](https://wolski.github.io/prolfquabenchmark/articles/Benchmark_proDA_medpolish.html) 
- [Benchmarking of msqrob2  using the IonStar dataset MaxQuant](https://wolski.github.io/prolfquabenchmark/articles/BenchmarkMSqRob2.html)  
- [Benchmarking of proDA and proflqua using IonStar dataset MSFragger IonQuant combined_protein.txt](https://wolski.github.io/prolfquabenchmark/articles/BenchmarkMSFraggerProteinIonStar.html)

# Estimating $A_{LOD}$

The main text describes how we estimate the abundance at the detection limit $A_{LOD}$. 
Figure \@ref(fig:missignessHistogram) shows the distribution of the average abundance within a group for all proteins. The red density shows the distribution for those proteins without a missing value, and the olive, turquoise, and lilac density show for proteins with 1,2 and 3 missing values, respectively. We see that the proteins with three missing values have, on average lower abundances than those with no missing values. We are using the median of the Lila density (one observation in the group), as the $A_{LOD}$ estimate.

```{r missignessHistogram, fig.cap= "Distribution of the average abundance within group a and c for all proteins. Each density are color coded depending the number of missing observations per protein in a group.", echo=FALSE}

datadir <- file.path(find.package("prolfquadata") , "quantdata")
inputMQfile <-  file.path(datadir,
                          "MAXQuant_IonStar2018_PXD003881.zip")
inputAnnotation <- file.path(datadir, "annotation_Ionstar2018_PXD003881.xlsx")
mqdata <- list()

mqdata$data <- prolfqua::tidyMQ_Peptides(inputMQfile)
mqdata$config <- prolfqua::create_config_MQ_peptide()
annotation <- readxl::read_xlsx(inputAnnotation)
res <- prolfqua::add_annotation(
  mqdata$data,
  annotation,
  fileName = "raw.file"
)
mqdata$config$table$factors[["dilution."]] = "sample"
mqdata$config$table$factors[["run_Id"]] = "run_ID"
mqdata$config$table$factorDepth <- 1
mqdata$data <- prolfqua::setup_analysis(res, mqdata$config)
lfqdata <- prolfqua::LFQData$new(mqdata$data, mqdata$config)
lfqdata$data <- lfqdata$data |> 
  dplyr::filter(dilution. == "a" | dilution. =="c")
lfqdata$remove_small_intensities()
bb <- lfqdata$get_Plotter()
bb$missigness_histogram()
```

# The probabilities produced by ROPECA are not p-values

In the main text discuss that the properties of the Beta distribution-based probabilities, computed from peptide level p-values, using the ROPECA method are not well understood. 
We are running here a simulation experiment, which shows that if we start with a dataset where the null hypothesis is valid for all the peptides, i.e., the p-values are uniformly distributed, the produced Beta distribution-based probabilities are not uniformely distributed.

For all the peptides (`nrPep = 10000`) $H0$ is true. We know that if $H0$ is true, the distribution of the $p-values$ will be uniform. (Figure \@ref(fig:figROPECA) left Panel). Therefore, we can omit the step of generating the data from $H0$ and computing the p-values but simulate the p-values directly by sampling from the uniform distribution (`runif(nrPep)`).

By sampling 800 protein id's 10000 times (with replacement), we assign the peptides to proteins ('sample(1:800, size = nrPep,
   replace = TRUE, ...)'). To make the data more realistic, that is that some proteins have a lot of peptides, and many have just a few peptides; the density of sampling the protein Id is exponential (`dexp(seq(0,5,length = 800)`). Figure \@ref(fig:figROPECA) center panel shows the number of proteins as a function of the number of peptides per protein. Afterward, we apply the ROPECA methods as discussed in the methods section of the manuscript. Figure \@ref(fig:figROPECA) shows the distribution of the Beta distribution-based probabilities, which is not uniform. Hence, they do not have the same interpretation as p-values, i.e., the probability of falsely rejecting the $H0$ if $H0$ is true.

```{r figROPECA , fig.cap = "Left panel - distibution of peptide level p.values. Center panel - number of proteins as a function of the number of peptides per protein. Right panel - distribution of protein level p-values(?) obtained with the ropeca method."}
set.seed(10)
nrPep <- 10000
nrProtein <- 800
p.value <- runif(nrPep)
estimate <- rnorm(nrPep)
avgAbd <- runif(nrPep)
protein_Id <- sample(1:800, size = nrPep,
   replace = TRUE, prob = dexp(seq(0,5,length = 800)))

testdata <- data.frame(contrast = "contrast1",
   protein_Id = protein_Id,
   estimate = estimate,
   pseudo_estimate = estimate,
   p.value = p.value,
   avgAbd = avgAbd )

xx30 <- prolfqua::summary_ROPECA_median_p.scaled(testdata,
                                     subject_Id = "protein_Id",
                                     estimate = "estimate",
                                     p.value = "p.value",
                                     max.n = 30)

par(mfrow = c(1,3))
hist(testdata$p.value, breaks = 20, xlab = "p-value", main = "")
plot(table(table(protein_Id)), xlab = "nr of peptides/protein", ylab = "# nr of proteins", main = "")
hist(xx30$beta.based.significance, breaks = 20, xlab = "p-value(?) obtained with Ropeca.", main = "")

```

# Creating a prolfqua configuration

The following code demonstrates how we use `r BiocStyle::Githubpkg("fgcz/prolfqua")` to analyze protein intensities reported in the MSFragger `combined_protein.tsv` file.
First, we create a tidy table containing the protein abundances by reading the `combined_protein.tsv` file using  `tidy_MSFragger_combined_protein.` Then, we read the sample annotation from the file `annotation.xlsx` file. Next, we create an `AnalysisTableAnnotation` R6 object.
Bottom-up proteomics data is hierarchical, i.e., a protein has peptides, peptides might be modified, etc. Therefore, the `AnalysisTableAnnotation` has a `hierarchy` field storing a list with an entry for each hierarchy level.
Since `combined_portein.tsv` only holds protein level data, the hierarchy list has one element, and we use it to specify which column contains the protein identifiers. We also need to define which column contains the protein abundances we want to use for the data analysis.
Finally, we have to specify which columns contain the explanatory variables of the analysis. The `AnalysisTableAnnotation` has the field `factors,` a list with as many entries as explanatory variables. Here we include two explanatory variables, the dilution, specified in the column 'sample', and 'run' stored in the column 'run_ID', representing the order of the measurement.

```{r echo=TRUE}
datadir <- file.path(find.package("prolfquadata") , "quantdata")
inputFragfile <-  file.path(datadir, "MSFragger_IonStar2018_PXD003881.zip")
inputAnnotation <- file.path(datadir, "annotation_Ionstar2018_PXD003881.xlsx")
# read input annotation
annotation <- readxl::read_xlsx(inputAnnotation)

protein <- tibble::as_tibble(
    read.csv(unz(inputFragfile,"IonstarWithMSFragger/combined_protein.tsv"),
             header = TRUE, sep = "\t", stringsAsFactors = FALSE))

# read combined_protein.tsv 
protein <- prolfqua::tidy_FragPipe_combined_protein_deprec(protein)
# remove proteins identified by a single peptide
protein <- protein |> 
  dplyr::filter(unique.stripped.peptides > 1)

# annotate the data
merged <- dplyr::inner_join(annotation, protein)
atable <- prolfqua::AnalysisTableAnnotation$new()
atable$fileName = "raw.file"
# specify column containing protein identifiers
atable$hierarchy[["protein_Id"]] = "protein"

# column with protein abundances
atable$set_response("total.intensity")

# the factors of the analysis
atable$factors[["dilution."]] = "sample"
atable$factors[["run"]] = "run_ID"

config <- prolfqua::AnalysisConfiguration$new(atable)

adata <- prolfqua::setup_analysis(merged, config)
lfqdata <- prolfqua::LFQData$new(adata, config)
# show number of proteins in the dataset
lfqdata$hierarchy_counts()

```

# Miscellaneous

(ref:tabCompletion) The screenshot displays the command-line completion (tab completion) of RStudio on the `prolfqua::LFQData` R6 object. In the example, it shows the getter methods of the object.

```{r tabCompletion, echo=FALSE, fig.cap="(ref:tabCompletion)", out.width = '66%'}
knitr::include_graphics("graphics/codeSnippet1TabCompletion.png")
```

```{r proLFQuaSticker, echo=TRUE, out.height="5cm", eval=TRUE, fig.cap="Sticker maintainer: Witold E. Wolski; License: Creative Commons Attribution CC-BY. Feel free to share and adapt, but don't forget to credit the author."}
file.path("graphics/hexStickerProlfqua.png") |>
  knitr::include_graphics()
```

# Session info {.unnumbered}

```{r sessionInfo, echo=FALSE}
pander::pander(sessionInfo())
```