### 2024-06-28 inst/triqler/Benchmark_Triqler.Rmd

---
title: "Looking into Triqler results"
output: html_document
date: "2023-04-18"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

## R Markdown

```{r}
library(readr)
library(tidyverse)
resTriq <- read_tsv("../inst/triqler/proteins.tsv")
resTriq$peptides <- NULL

resTriq <- as.data.frame(resTriq)
resTriq$DECOY <- ifelse(grepl("^DECOY_", resTriq$protein), TRUE, FALSE)
resTriq$protein <-  gsub("^DECOY_", "", resTriq$protein)
resTriq <- resTriq |> tidyr::separate(protein, into =  c("ID", "Organism"), sep = "_",remove = FALSE)
resTriq$Organism |> table()

resTriq <- resTriq |> mutate(color = case_when(Organism == "YEAST" ~ "green",
                                               Organism == "HUMAN" ~ "black",
                                               Organism == "ECOLI" ~ "green",
                                               TRUE ~ "pink"))

resTriq <- resTriq |> arrange(color)
plot(resTriq$log2_fold_change, -log10(resTriq$q_value), col = resTriq$color, pch = 20)
Honly <- resTriq |> filter(color == "black")
points(Honly$log2_fold_change, -log10(Honly$q_value), pch = 20)

abline(h = -log10(0.1))
abline(v = c(-1,1))

```

# Analyse data with prolfqua

```{r}
res <- read_tsv("../inst/triqler/LFQBench_PS.csv")
res$condition <- paste0("c",res$condition)
res <- res |> filter(!grepl("^DECOY_",res$proteins))
hist(res$searchScore)

# res <- res |> filter(searchScore > 9)

resAgr <- res |> group_by(run, condition, peptide, proteins ) |>
  summarize(intensity = sum(intensity, na.rm = TRUE), searchScore = max(searchScore, na.rm = TRUE), n = n())

resAgr <- ungroup(resAgr)
resAgr$identQ <- max(resAgr$searchScore) - resAgr$searchScore
resAgr <- resAgr |> filter(!grepl("^DECOY_",resAgr$proteins))

atable <- prolfqua::AnalysisTableAnnotation$new()
atable$hierarchy[["protein_Id"]] <- "proteins"
atable$hierarchy[["peptide_Id"]] <- "peptide"
atable$hierarchy
atable$hierarchy_keys()
atable$factors[["condition"]] <- "condition"
atable$ident_Score <- "searchScore"
atable$ident_qValue <- "identQ"
atable$set_response("intensity")
atable$fileName <- "run"
config <- prolfqua::AnalysisConfiguration$new(atable)

undebug(prolfqua::setup_analysis)
xx <- prolfqua::setup_analysis(resAgr, config)

library(prolfqua)
psmdata <- LFQData$new(xx, config)
psmdata$to_wide()$data |> dim()

psmdata$config$table$hierarchyDepth <- 1
psmdata$config$table$hierarchy_keys_depth()
psmdata$hierarchy_counts()
psmdata$remove_small_intensities()

tr <- psmdata$get_Transformer()$intensity_array(log)
ag <- tr$lfq$get_Aggregator()
ag$medpolish()
protd <- ag$lfq_agg
tr <- protd$get_Transformer()
tr$intensity_array(exp, force = TRUE)
protd <- tr$lfq
protd$is_transformed(FALSE)

stprot <- protd$get_Stats()
stprot$violin_median()

pl <- protd$get_Plotter()
hm <- pl$heatmap()

```

```{r}
print(hm)

```

```{r}

pl$intensity_distribution_violin()
tr <- protd$get_Transformer()
tr$intensity_matrix(vsn::justvsn)
lfqdata <- tr$lfq
lfqdata$get_Plotter()$intensity_distribution_violin()
hm <- lfqdata$get_Plotter()$heatmap()
```

```{r}
print(hm)
```

```{r}
st <- lfqdata$get_Stats()
st$violin()

```

## Prolfqua contrasts

```{r}
lfqdata$rename_response("intensity")
contr <- c("2vs1" = "conditionc2 - conditionc1")
lmstr <- strategy_lm("intensity ~ condition")
models <- build_model(lfqdata, lmstr)
sum(models$modelDF$isSingular,na.rm = TRUE)
sum(is.na(models$modelDF$isSingular))

contrlm <- Contrasts$new(models, contr)
contrmiss <- ContrastsMissing$new(lfqdata, contr) 
contrlmmod <- contrlm |> ContrastsModerated$new()
contrmissmod <- contrmiss |> ContrastsModerated$new()

contrmergmod <- prolfqua::merge_contrasts_results(contrlmmod, contrmissmod)$merged
ctrmergmod <- contrmergmod$get_contrasts()

```

```{r}

pl <- contrmergmod$get_Plotter()
pl$contrastDF$color <- ifelse(grepl("HUMAN",ctrmergmod$protein_Id), "black", "red")
pl$volcano(colour = "color")

```

```{r}

tripleprot_bench_preprocess <- function(data, idcol = "protein_Id", drop = "YEAST") {

  truepos = setdiff(c("YEAST", "ECOLI"), drop)
  print(truepos)

  tmp <- data |>
    ungroup() |>
    mutate(species  = case_when(
      grepl("HUMAN", !!sym(idcol)) ~ "HUMAN",
      grepl("ECOLI", !!sym(idcol)) ~ "ECOLI",
      grepl("YEAST", !!sym(idcol)) ~ "YEAST",
      TRUE ~ "OTHER"
    ))
  res <- tmp |> dplyr::filter(!.data$species == "OTHER"  & !.data$species == drop)
  res <- res |> mutate(TP = (.data$species == truepos))
  return(list(data = res , table = table(tmp$species)))
}

```

```{r moderated}
ttd <- tripleprot_bench_preprocess(ctrmerg,drop = "YEAST" )
dim(ttd$data)
benchmark_mod_prolfqua <- prolfqua::make_benchmark(
  ttd$data,
  model_description = "prolfqua_merged",
  model_name = "prolfqua_merged",
  FDRvsFDP = list(list(score = "FDR", desc = FALSE))
)
benchmark_mod_prolfqua$pAUC_summaries()$ftable$content
knitr::kable(benchmark_mod_prolfqua$pAUC_summaries()$ftable$content)
prolfqua::table_facade(benchmark_mod_prolfqua$smc$summary, "Nr of estimated contrasts")
benchmark_mod_prolfqua$plot_score_distribution()
benchmark_mod_prolfqua$plot_ROC(0.05)
benchmark_mod_prolfqua$plot_FDRvsFDP()

```

```{r}
ttd <- tripleprot_bench_preprocess(ctrmergmod,drop = "ECOLI" )
dim(ttd$data)
benchmark_mod_prolfqua <- prolfqua::make_benchmark(
  ttd$data,
  model_description = "prolfqua_merged",
  model_name = "prolfqua_merged",
  FDRvsFDP = list(list(score = "FDR", desc = FALSE)),
  benchmark = list(
    list(score = "diff", desc = FALSE),
    list(score = "statistic", desc = FALSE),
    list(score = "scaled.p.value", desc = FALSE)
  )
)

benchmark_mod_prolfqua$pAUC_summaries()$ftable$content
prolfqua::table_facade(benchmark_mod_prolfqua$smc$summary, "Nr of estimated contrasts")
benchmark_mod_prolfqua$plot_score_distribution()
benchmark_mod_prolfqua$plot_ROC(0.05)

benchmark_mod_prolfqua$plot_FDRvsFDP()
benchmark_mod_prolfqua$.get_confusion
prolfqua:::ms_bench_add_scores

dx <- benchmark_mod_prolfqua$get_confusion_FDRvsFDP()
head(dx)
plot(dx$scorecol, dx$FDP_)

```

```{r}
resTriqM <- resTriq |> filter(DECOY == FALSE)
nrow(resTriqM)

resTriqM <- resTriqM |> select(protein,log2_fold_change,q_value, color)
ctr <- ungroup(ctrmergmod) |> select(protein_Id,diff,FDR)

xx <- inner_join(resTriqM, ctr, by = c(protein = "protein_Id"))

plot( xx$log2_fold_change, -xx$diff, col = xx$color)
abline(0, 1, col = "red", lwd = 2)
abline(h = c(-1,1), v = c(-1,1))

plot(-log10(xx$FDR), -log10(xx$q_value), log="xy", col = xx$color )
abline(h = -log10(0.05), v = -log10(0.05))
```

```{r}

```